---
title: "Accidents In Great Britain(2005-2015)"
author: "Mahendra Kumar lal"
date: "February 25, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, echo = FALSE, 
                      message = FALSE, warning = FALSE)
```


## Introduction

- This project was created as part of the Developing Data Products Week 3; a course of the Coursera [Data Science Specialisation](https://www.coursera.org/specializations/jhu-data-science).

- The goal of the project is to create a web page presentation using R Markdown that features a plot created with Plotly, and to host the resulting web page on either GitHub Pages, RPubs, or NeoCities.

- The interactive plot on the next slide represents the number of road accidents in Great Britain from 2005 to 2015, grouped by severity (slight, serious, or fatal).

    + A Loess smoother line has been added to highlight the overall evolution of the number of accidents.

```{r }

rm(list=ls())

library(plotly)
library(data.table)
library(tidyr)
library(lubridate)
library(zoo)

```

```{r load_data, results='hide'}
#Read Data for 2005-20014
URL<-"http://data.dft.gov.uk.s3.amazonaws.com/road-accidents-safety-data/Stats19_Data_2005-2014.zip"
temp<-tempfile()
download.file(URL, temp)
accident01<-read.csv(unz(temp,"Accidents0514.csv"),na = c("","NA"), blank.lines.skip = TRUE,header = TRUE)
unlink(temp)
str(accident01)

#READ Data for 2015
URL<-"http://data.dft.gov.uk/road-accidents-safety-data/RoadSafetyData_Accidents_2015.zip"
temp<-tempfile()
download.file(URL, temp)
accident02 <- read.csv(unz(temp,"Accidents_2015.csv"),na = c("","NA"), blank.lines.skip = TRUE,header = TRUE)
unlink(temp)
str(accident02)
#Keep Data of Severity and Date column
accident01 <- accident01 %>% select(Accident_Severity, Date)
accident02 <- accident02 %>% select(Accident_Severity, Date)

# Merge both the table and free up environment
accidents <- rbind(accident01, accident02)
rm(list = c("accident01", "accident02"))
str(accidents)

# Process the Data
# Convert severity into factor and add labels
accidents$Accident_Severity <-  factor(accidents$Accident_Severity, levels = 1:3, labels = c("Fatal", "Serious", "Slight"))
# convert date strings to Date objects
accidents$Date <- dmy(accidents$Date)
# group data by date and severity, get count group by date
accident_count <- accidents %>% 
  group_by(Date, Accident_Severity) %>%
  summarise(count = n()) %>%
  spread(key = Accident_Severity, value = count) %>% 
  as.data.frame()
# create a smoother for each severity to visualise general trends
loess_slight <- loess(Slight ~ as.numeric(Date), 
                      data = accident_count)
loess_serious <- loess(Serious ~ as.numeric(Date), 
                       data = accident_count)
loess_fatal <- loess(Fatal ~ as.numeric(Date), 
                     data = accident_count)
```

## Road accidents in GB (2005-2015)

```{r plot}
# plot data

plot_ly(accident_count) %>%
  add_trace(x = ~Date, y = ~Slight, type="scatter", mode = "markers", 
            name = "slight", legendgroup = "slight", 
            marker = list(color = "#52A9BD")) %>%
  add_trace(x = ~Date, y = ~Serious, type="scatter", mode = "markers",
            name = "serious", legendgroup = "serious", 
            marker = list(color = "#FFF16B")) %>%
  add_trace(x = ~Date, y = ~Fatal, type="scatter", mode = "markers",
            name = "fatal", legendgroup = "fatal", 
            marker = list(color = "#F5677D")) %>%
  add_trace(x = as.Date(loess_slight$x), y = fitted(loess_slight),
            type="scatter", mode = "lines",
            line = list(color = '#1A7A90'), 
            name = "slight Loess smoother", legendgroup = "slight", 
            hoverinfo = 'none', showlegend = FALSE) %>%
  add_trace(x = as.Date(loess_serious$x), y = fitted(loess_serious),
            type="scatter", mode = "lines",
            line = list(color = '#E9D625'),
            name = "serious Loess smoother", legendgroup = "serious",
            hoverinfo = 'none', showlegend = FALSE) %>%
  add_lines(x = as.Date(loess_fatal$x), y = fitted(loess_fatal),
            type="scatter", mode = "lines",
            line = list(color = '#DC2340'),
            name = "fatal Loess smoother", legendgroup = "fatal",
            hoverinfo = 'none', showlegend = FALSE) %>%
  layout(
    xaxis = list(title = "date"),
    yaxis = list(title = "number of accidents")
  )
```

## Notes

### Source code

- The R Markdown source code of this presentation can be found in [this GitHub repository](https://github.com/mahenlal/CourseraDevelopingDataProducts).

### Data source and licence

- This project uses the [road safety data set](https://data.gov.uk/dataset/road-accidents-safety-data), released by the United Kingdom's Department for Transport under the [Open Government Licence](http://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/).